Timeliness of security updates for Android

### 2014-11-10 securityupdates.tex
add some missing references

We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, \texttt{\url{http://androidvulnerabilities.org/}}.
We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
We have received submissions or amendments from \avoNumSubmitters\ individuals.
\avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular manufacturers and we know enough information about the vulnerability to tie it to the particular versions it affects.
We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO} and summarise them in Table~\ref{tab:andvulns}.

@@

\section{Discussion}
There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
When Capsicum is included in Linux~\cite{TODO} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming critical vulnerabilities.

@@

Examining security updates has been a key way of assessing security for many years.
Many software update systems have been found not to authenticate the connection to the update server and/or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
Even package management systems designed to provide secure updates have been found to have vulnerabilities~\cite{Cappos2008}.
Fortunately Android does authenticate the update binaries (though the APK vulnerabilities circumvented this) and Google Play downloads them over a secure connection~\cite{Viennot2014}. \todolater{double check that}

@@

User-Agent strings have been used to investigate how web browser versions change with updates~\cite{Frei2008} and found that at most 80\% of Firefox users were running the most recent version.
The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
Android's update process is manual, the user is notified but must action the upgrade which will then download the update and install it, the phone is rendered inoperable during the update process which is inconvenient for users and the phone must have sufficient charge so that it will not run out during the update process.
Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
The same technique would be more difficult on phones with limited storage space (as for many cheap Android phones which have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
Google is deploying the same silent update technique through Google Play Services~\cite{TODO} which automatically installs updates for core Google components of Android, this also bypasses the manufacturer and operator.

>>>>

We developed and maintain an open platform for filing critical vulnerabilities in a machine readable format, \texttt{\url{http://androidvulnerabilities.org/}}.
We seeded it with data from the CVE database, vendor lists, reports from the literature and various forums.
We have received submissions or amendments from \avoNumSubmitters\ individuals.
\avo\ contains \avoNumVulnerabilities\ vulnerabilities of which \avoNumVulnAllAndroid\ affect all Android devices and \avoNumVulnSpecific\ are specific to particular devices or manufacturers.
We only use the \daNumVulnsUsed\ vulnerabilities which we have sufficient data to use for this analysis because they affect all Android devices rather than particular manufacturers and we know enough information about the vulnerability to tie it to the particular versions it affects.
We summarise the \daNumVulnsUsed\ vulnerabilities used for this analysis in Table~\ref{tab:andvulns}.\todolater{We have published full details of the \daNumVulnsUsed\ vulnerabilities used for this analysis in an accompanying technical report~\cite{TODO}}

@@

\section{Discussion}
There are continuing efforts to reduce the impact of critical vulnerabilities, both in Android and more widely.
SEAndroid~\cite{Smalley2013} which is included in Android from version 4.1~\cite{jelly-bean-release} claimed to prevent some root vulnerabilities and to reduce the impact of others.
Capability based enforcement systems such as Capsicum~\cite{Watson2010} substantially reduce the capabilities that an exploit has to try and gain increased privilege with.
When Capsicum is included in Linux\footnote{\url{https://github.com/google/capsicum-linux}} and hence in Android, it could be used to place fine grained restrictions on system daemons, preventing vulnerabilities becoming critical vulnerabilities.

@@

Examining security updates has been a key way of assessing security for many years.
Many software update systems have been found not to authenticate the connection to the update server and/or do not authenticate the downloaded binaries~\cite{Bellissimo2006}.
Even package management systems designed to provide secure updates have been found to have vulnerabilities~\cite{Cappos2008}.
Fortunately Android does authenticate the update binaries (though the APK vulnerabilities circumvented this) and Google Play downloads them over a secure connection~\cite{Viennot2014}.

@@

User-Agent strings have been used to investigate how web browser versions change with updates~\cite{Frei2008} and found that at most 80\% of Firefox users were running the most recent version.
The same analysis was used to show that Chrome's use of silent updates seems to increase uptake of upgrades~\cite{Duebendorfer2010} with 97\% of users running the latest version within 3 weeks of release.
Android's update process is manual, the user is notified but must action the upgrade which will then download the update and install it, the phone is rendered inoperable during the update process which is inconvenient for users and the phone must have sufficient charge so that it will not run out during the update process.
Partly this is the result of the fact that an operating system update is being installed and so a reboot is required, but Chrome installs the new version side by side with the old one and switches the next time it is restarted.
The same technique would be more difficult on phones with limited storage space (as for many cheap Android phones which have barely enough space to install just the update) but is a plausible improvement for more high-end devices.
Google is deploying the same silent update technique through Google Play Services\footnote{\url{http://lifehacker.com/why-google-play-services-are-now-more-important-than-an-975970197}} which automatically installs updates for core Google components of Android, this also bypasses the manufacturer and operator.
